---
output: pdf_document
---
Getting and cleaning Data:Project
======================================
Created by Krish Mahajan 

### Basic settings
```{r}
echo = TRUE  # Always make code visible
options(scipen = 1)  # Turn off scientific notations for numbers
```

### Merges the training and the test sets to create one data set.
```{r merge,cache=TRUE}
trainData <- read.table("./data/train/X_train.txt")
dim(trainData)
head(trainData,n=3)

trainLabel <- read.table("./data/train/y_train.txt")
table(trainLabel)

trainSubject <- read.table("./data/train/subject_train.txt")
head(trainSubject,n=3)
dim(trainSubject)

#test data
testData <- read.table("./data/test/X_test.txt")
dim(testData)

testLabel <- read.table("./data/test/y_test.txt") 
table(testLabel) 

testSubject <- read.table("./data/test/subject_test.txt")

### Joining test & training Data
joinData <- rbind(trainData, testData)
dim(joinData) 

joinLabel <- rbind(trainLabel, testLabel)
dim(joinLabel) 

joinSubject <- rbind(trainSubject, testSubject)
dim(joinSubject) 
```

### Step2. Extracts only the measurements on the mean and standard 


```{r extraction,cache=TRUE}
 
features <- read.table("./data/features.txt")
dim(features) 
head(features)
meanStdIndices <- grep("mean\\(\\)|std\\(\\)", features[, 2])
length(meanStdIndices)
joinData <- joinData[, meanStdIndices]
dim(joinData) 
names(joinData) <- gsub("\\(\\)", "", features[meanStdIndices, 2]) # remove "()"
names(joinData) <- gsub("mean", "Mean", names(joinData)) # capitalize M
names(joinData) <- gsub("std", "Std", names(joinData)) # capitalize S
names(joinData) <- gsub("-", "", names(joinData)) # remove "-" in column names 
names(joinData)
```

### Step3. Uses descriptive activity names to name the activities in 

```{r descriptive,cache=TRUE}

activity <- read.table("./data/activity_labels.txt")
head(activity)
dim(activity)

activity[, 2] <- tolower(gsub("_", "", activity[, 2])) # remove '-' and lower 
substr(activity[2, 2], 8, 8) <- toupper(substr(activity[2, 2], 8, 8)) #codingConventions
substr(activity[3, 2], 8, 8) <- toupper(substr(activity[3, 2], 8, 8)) #codingConventions

activityLabel <- activity[joinLabel[, 1], 2] #interesting step
joinLabel[, 1] <- activityLabel
names(joinLabel) <- "activity"
```

### Step4. Appropriately labels the data set with descriptive activity 


```{r descriptive ,cache=TRUE}

names(joinSubject) <- "subject"
cleanedData <- cbind(joinSubject, joinLabel, joinData)
dim(cleanedData)
write.table(cleanedData, "merged_data.txt") # write out the 1st dataset
```


### Step5. Creates a second, independent tidy data set with the average of  each variable for each activity and each subject.

```{r tidyDataset,cache=TRUE}
subjectLen <- length(table(joinSubject))
activityLen <- dim(activity)[1] 
columnLen <- dim(cleanedData)[2]
result <- matrix(NA, nrow=subjectLen*activityLen, ncol=columnLen) 
result <- as.data.frame(result)
colnames(result) <- colnames(cleanedData)
row <- 1
for(i in 1:subjectLen) {
    for(j in 1:activityLen) {
        result[row, 1] <- sort(unique(joinSubject)[, 1])[i]
        result[row, 2] <- activity[j, 2]
        bool1 <- i == cleanedData$subject
        bool2 <- activity[j, 2] == cleanedData$activity
        result[row, 3:columnLen] <- colMeans(cleanedData[bool1&bool2, 3:columnLen])
        row <- row + 1
    }
}
head(result)
write.table(result, "data_with_means.txt") # write out the 2nd dataset
```
